Estimating the Initial Susceptible Fraction in Dengue Dynamics

### 2015-01-27 draft_lm1.tex
updated table 1 an improved results and discussion

Figure \ref{Fig:S0}, shows the model from (\ref{eq:model}) fitted to the data. 
In it we can see that the susceptibles series in each epidemic starts at the 
estimated level of $S_0$.
The proportion of susceptibles may seem low, but we 
must remember that these estimates are being affected by an unknown 
underreporting factor, which experts suggest is somewhere between 5 and 10, 
i.e. for every case observed there are 5 or 10 unobserved.
Since this 
underreporting ration affects both the numerator and denominator of 
(\ref{eq:AR}), its effects should cancel out, giving us an unbiased attack ratio 
estimate.
One other possible source of bias which would lead to the 
underestimation of $S_0$ could come from a significant part of the population 
not being exposed to the disease.
However, as we can see in figure \ref{fig:mapas} , the disease spreads pretty 
much homogeneously over 
the entire city (in the last four epidemics, at least).

@@

Table \ref{tab:AR} contains the attack ratios calculated for each 
epidemic/outbreak. It is interesting to notice that the larger epidemics, in 
terms of total number of cases are not the one with the greater attack ratios. 
% We only support three levels of headings, please do not create a heading level below \subsubsection.
%\subsection*{Subsection 1}
\section*{Discussion}
% Do NOT remove this, even if you are not including acknowledgments.
\section*{Acknowledgments}
LMC is grateful to Dr. Leonardo Bastos for useful discussions on the 
posterior inference for $R(t)$.
The authors are also grateful to Claudia T. 
Codeço for helpful discussions about the manuscript.

@@

\begin{table}[!ht]
\caption{
\bf{Median Attack ration calculated according to (\ref{eq:AR2})}. $^\dag$: Year 
corresponds to the start of the epidemic. Sometimes the peak of cases occur in 
the following year.}
\begin{center}
\begin{tabular}{l|r}
\hline
Year$^\dag$ & Median Attack Ratio (\%) \\
\hline
1996 & \textbf{38.64} (17.18-54.19)\\
1997 & \textbf{87.31} (73.83-87.31)\\
1998 & \textbf{49.86} (49.19-49.86)\\
1999 & \textbf{10.76} (3.72-20.24)\\
2000 & \textbf{25.49} (24.47-26.98)\\
2001 & \textbf{48.12} (46.57-49.15)\\
2005 & \textbf{14.74} (10.23-21.38)\\
2006 & \textbf{11.22} (8.00-14.41)\\
2007 & \textbf{15.08} (14.76-15.38)\\
2008 & \textbf{13.73} (3.10-30.84)\\
2010 & \textbf{18.30} (17.34-19.33)\\
2011 & \textbf{8.59} (8.20-9.36)\\
2012 & \textbf{13.92} (12.71-14.92)\\
\hline
\end{tabular}
\begin{flushleft} We report medians and 95\% credibility intervals.
\end{flushleft}

>>>>

Figure \ref{Fig:S0}, shows the model from (\ref{eq:model}) fitted to the data. 
In it we can see that the susceptibles series in each epidemic starts at the 
estimated level of $S_0$.
The proportion of susceptibles may seem low, but we 
must remember that these estimates are being affected by an unknown 
underreporting factor, which experts suggest is somewhere between 5 and 10, 
i.e. for every case observed there are 5 or 10 unobserved.
Since this underreporting affects both the numerator and denominator of 
(\ref{eq:AR}), its effects should cancel out, giving us an unbiased attack ratio 
estimate.
One other possible source of bias which would lead to the 
underestimation of $S_0$ could come from a significant part of the population 
not being exposed to the disease.
However, as we can see in figure \ref{fig:mapas} , the disease spreads pretty 
much homogeneously over 
the entire city (in the last four epidemics, at least).

@@

Table \ref{tab:AR} contains the attack ratios and medians of the $S_0$ 
estimated for each epidemic/outbreak. It is interesting to notice that the 
larger epidemics, in terms of total number of cases are not the one with the 
greater attack ratios. This stresses the importance of knowing the imunological 
structure of the population in order to truly assess the impact of eny 
epidemic.

The authors hope that this evidence will motivate public health 
outhorities to invest in annual serological surveys, to determine the profile 
of susceptibility to each dengue virus as well as to estimate the 
underreporting factor of the notification system.
% We only support three levels of headings, please do not create a heading level below \subsubsection.
%\subsection*{Subsection 1}
%\section*{Discussion}
% Do NOT remove this, even if you are not including acknowledgments.
\section*{Acknowledgments}
LMC is grateful to Dr. Leonardo Bastos for useful discussions on the 
posterior inference for $R(t)$.
The authors are also grateful to Claudia T. 
Code\c{c}o for helpful discussions about the manuscript.

@@

\begin{table}[!ht]
\caption{
\bf{Median Attack ratio and 95\% credibility intervals calculated according to 
(\ref{eq:AR2})}. $^\dag$: Year corresponds to the start of the epidemic. 
Sometimes the peak of cases occur in the following year.}
\begin{center}
\begin{tabular}{c|c|c}
\hline
Year$^\dag$ & median Attack Ratio & $S_0$ \\
\hline
1996 & 38.64\% (17.18\%-54.19\%) & 0.171\%\\
1997 & 87.31\% (73.83\%-87.31\%) & 0.273\%\\
1998 & 49.86\% (49.19\%-49.86\%) & 0.142\%\\
1999 & 10.76\% (3.72\%-20.24\%) & 0.345\%\\
2000 & 25.49\% (24.47\%-26.98\%) & 1.547\%\\
2001 & 48.12\% (46.57\%-49.15\%) & 4.951\%\\
2005 & 14.74\% (10.23\%-21.38\%) & 1.472\%\\
2006 & 11.22\% (8.00\%-14.41\%) & 2.810\%\\
2007 & 15.08\% (14.76\%-15.38\%) & 13.451\%\\
2008 & 13.73\% (3.10\%-30.84\%) & 0.672\%\\
2010 & 18.30\% (17.34\%-19.33\%) & 4.540\%\\
2011 & 8.59\% (8.20\%-9.36\%) & 21.482\%\\
2012 & 13.92\% (12.71\%-14.92\%) & 6.208\%\\

\hline
\end{tabular}
